---
title: "Road Safety in the United Kingdom in 2015"
author: "André Diegues, André Fonseca, Rui Andrade"
date: "17 de Março de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Association rule mining is a concept that fits under descriptive tasks in the area of _Data Mining_. Originally used in what is called _Market Basket Analysis_, so as to analyse customer purchases and try to find unexpected associations between items, it has been generalised to help explain, for example, what factors could contribute towards a given characteristic in an observation. Sets of items that are commonly found together are called _frequent itemsets_ and relationships between items - in the sense that having certain items could favour having other certain items - are called _association rules_.

In this report, we study a dataset that contains several details on car accidents in the United Kingdom. This dataset was made available by the Government of the United Kingdom and dates back to 2015. Examples of information contained in it include: GPS location of the occurence, severity of accident, number of vehicles and casualties involved, time and date, type of road, speed limit, natural conditions (weather or light), road conditions and infrastructure available in the road (how are junctions handled, are there crossings available or not...). The goal of this work is to find frequent itemsets relating to the details of accidents, or in other words, what do accidents typically have related to one another, as well as finding association rules on cause and effect of traffic accidents.

```{r cars}
summary(cars)
```

## Preprocessing
We have done some preprocessing on the data in order to find interesting relations using association rules. We start by creating a function that allowed us to read the xls file guides for our main dataset and consequently the xls file with the main dataset.
This allowed us to exlude the second and third columns , also the Thirty first column as well all the variables relative to local authority as we found them not correlated with the rest of the set. 
We also verified that the dataset had missing values and therefore we excluded these observations from the data.
In a first analysis of the dataset we realized that for a better perception of the data we should divide some of the variables to Nominal variables intervals , only to improve our goal to find association rules on the cause and effect of traffic  accidents. In order to be more precise , we are going to make a exploratory analysis an then we may proceed on excluding the variables that we do not think that were not necessary in our analysis.


## Exploratory Analysis
For this part, we started by matching some of the code guides with our actual data for better interpretation.
After that we start by generating some graphs with a certain interest to visualize some correlation.
For that porpuse our first plot as we can see in _fig.1_ tries to find association between accident locations and the severity of the accident.

``` associar aquele fig 1 ao plot```
```{r pressure, echo=FALSE}
library(ggplot2)

# scatterplot of accident locations, color coded by level of severity
ggplot(data, aes(x=Longitude, y=Latitude, colour=Accident_Severity)) + 
  geom_point() + 
  ggtitle('Accident locations and level of severity')
```

As we can see we have scatter the accident locations per level of severity , and found that the geografical dispersion is similir to the UK geographical terrain, and the most frequent cases of Fatal accident (code = 1) are in the north side of the country, looking over this points it is quite reasonable to assume that fatal accidents are directly related to the worst atmospheric conditions that are felt normaly with more intensity in the north of the country (during winter).

Now our goal was to relate to each type of accident severity with a month, and know in this way the distribution of accidents per moth as we can see in _fig.2_.

```{r pressure, echo=FALSE}
# severity of accidents by month
data_graph2 <- dplyr::select(data, c(Month, Accident_Severity))
data_graph2 <- dplyr::group_by(data_graph2, Accident_Severity)
data_graph2 <- dplyr::count(data_graph2, Month, Accident_Severity)
ggplot(data_graph2, aes(y=n, x=Accident_Severity)) + 
  geom_bar(stat="identity") + 
  ggtitle('Severity of accidents by month') + facet_wrap(~ Month)
```

Looking to _fig.2_ we immediately discovered that we have a decreasing distribution of accident from Slight (code=3) to fatal(code=1) almost identical on all year.
Observing the data in a more profound way, we observe that the month of December is the most deadly and in some way supports our claims on _fig.1_.

Now our goal was to find how the Accident Severity would behave when compared to the atmospheric conditions.
From this data we expect to see a higher frequency on Fatal accidents when demostrated agaist each type of weather conditions as we can see on _fig.3_.

```{r pressure, echo=FALSE}
# severity of accidents by weather conditions
data_graph3 <- dplyr::select(data, c(Weather_Conditions, Accident_Severity))
data_graph3 <- dplyr::group_by(data_graph3, Weather_Conditions, Accident_Severity)
data_graph3 <- dplyr::count(data_graph3, Weather_Conditions, Accident_Severity)
data_graph3 <- dplyr::mutate(data_graph3, fr=n/sum(n))
ggplot(data_graph3, aes(y=fr, x=Accident_Severity)) + 
  geom_bar(stat="identity") + 
  ggtitle('Severity of accidents by weather conditions') + facet_wrap(~ Weather_Conditions)
```

Observing our _fig.3_ we can perceive that is indeed when it's verified bad wheather conditions like rain, fog and high winds there's higher frequency on Fatal accidents.
We can assume that when Snowing drivers have limitions on their driving conditions and because of that the frequency of Fatal accidents is almost null when compared to other conditions.

We also found interesting knowing the distribution of casualties per accident, this data could provide some perspective 



You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
